---
title: "Regression code collection"
author: "Harold Trammel"
date: "September 10, 2015"
output: html_document
---
```{r setup, 
      cache = FALSE, 
      echo = TRUE, 
      message = FALSE, 
      warning = FALSE, 
      tidy = TRUE}

library(ggplot2)
library(UsingR)
data(diamond)
```

This document collects various code snippets from the Regression course by Brian Caffo.  Most of the code originates with him.  

I am collecting this code for learning and future reference.


### Professor Caffo's work with diamond data
## Example
### `diamond` data set from `UsingR` 
Data is diamond prices (Singapore dollars) and diamond weight
in carats (standard measure of diamond mass, 0.2 $g$). To get the data use `data(diamond)`


### Plot of the data
```{r plot_1}
g = ggplot(diamond, aes(x = carat, y = price))
g = g + xlab("Mass (carats)")
g = g + ylab("Price (SIN $)")
g = g + geom_point(size = 7, colour = "black", alpha=0.5)
g = g + geom_point(size = 5, colour = "blue", alpha=0.2)
g = g + geom_smooth(method = "lm", colour = "black")
print(g)
```

### Fitting the linear regression model
```{r}
fit <- lm(price ~ carat, data = diamond)
coef(fit)
```

* We estimate an expected `r round(coef(fit)[2], 2)` (SIN) dollar increase in price for every carat increase in mass of diamond.
* The intercept `r round(coef(fit)[1], 2)` is the expected price
  of a 0 carat diamond.

### Getting a more interpretable intercept
```{r}
fit2 <- lm(price ~ I(carat - mean(carat)), data = diamond)
coef(fit2)
```

Thus $`r round(coef(fit2)[1], 1)` is the expected price for 
the average sized diamond of the data (`r mean(diamond$carat)` carats).

```{r}
fit3 <- lm(price ~ I(carat * 10), data = diamond)
coef(fit3)
```


## Predicting the price of a diamond
```{r}
newx <- c(0.16, 0.27, 0.34)
coef(fit)[1] + coef(fit)[2] * newx
predict(fit, newdata = data.frame(carat = newx))
```


Predicted values at the observed Xs (red)
and at the new Xs (lines)
```{r plot_2}
data(diamond)
plot(diamond$carat, diamond$price,  
     xlab = "Mass (carats)", 
     ylab = "Price (SIN $)", 
     bg = "lightblue", 
     col = "black", cex = 1.1, pch = 21,frame = FALSE)
abline(fit, lwd = 2)
points(diamond$carat, predict(fit), pch = 19, col = "red")
lines(c(0.16, 0.16, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.16,
      coef(fit)[1] + coef(fit)[2] * 0.16))
lines(c(0.27, 0.27, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.27,
        coef(fit)[1] + coef(fit)[2] * 0.27))
lines(c(0.34, 0.34, 0.12), 
      c(200, coef(fit)[1] + coef(fit)[2] * 0.34,
        coef(fit)[1] + coef(fit)[2] * 0.34))
text(newx, rep(250, 3), labels = newx, pos = 2)
```

```{r doing_equations_hard_way}
y <- diamond$price; x <- diamond$carat; n <- length(y)
# slope calculations
beta1 <- cor(y, x) * sd(y) / sd(x)  # slope
beta0 <- mean(y) - beta1 * mean(x)  # intercept
# residuals
e <- y - beta0 - beta1 * x
# estimate of sigma
sigma <- sqrt(sum(e^2) / (n-2)) 
# sums of squares of x's
ssx <- sum((x - mean(x))^2)
# standard error for slope at intercept
seBeta0 <- (1 / n + mean(x) ^ 2 / ssx) ^ .5 * sigma 
# standard error for around regression line
seBeta1 <- sigma / sqrt(ssx)
# t-statistics
tBeta0 <- beta0 / seBeta0
tBeta1 <- beta1 / seBeta1
# p-values 
pBeta0 <- 2 * pt(abs(tBeta0), df = n - 2, lower.tail = FALSE)
pBeta1 <- 2 * pt(abs(tBeta1), df = n - 2, lower.tail = FALSE)


coefTable <- rbind(c(beta0, seBeta0, tBeta0, pBeta0), c(beta1, seBeta1, tBeta1, pBeta1))
colnames(coefTable) <- c("Estimate", "Std. Error", "t value", "P(>|t|)")
rownames(coefTable) <- c("(Intercept)", "x")

fit <- lm(y~x)
summary(fit)$coefficients

```

## Getting a confidence interval
```{r}
sumCoef <- summary(fit)$coefficients
sumCoef[1,1] + c(-1, 1) * qt(.975, df = fit$df) * sumCoef[1, 2]
(sumCoef[2,1] + c(-1, 1) * qt(.975, df = fit$df) * sumCoef[2, 2]) / 10
```
With 95% confidence, we estimate that a 0.1 carat increase in
diamond size results in a `r round((sumCoef[2,1] - qt(.975, df = fit$df) * sumCoef[2, 2]) / 10, 1)` to `r round((sumCoef[2,1] + qt(.975, df = fit$df) * sumCoef[2, 2]) / 10, 1)` increase in price in (Singapore) dollars.

